---
title: "my-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The linreg function fits a linear regression model using ordinary least squares.

### Step 1: Load the package
Start by loading the *smlm* package.
```{r setup}
library(bonuslab)
```


### Step 2: Load the data
We will use the R built-in data set iris.
```{r data}
data("iris")
head(iris)
```

### Step 3: Create a formula
Create a formula with a dependent variable and at least one independent variable. The following formula uses *Petal.Length* as the dependent variable, and *Species* as the independent variable.
```{r formula}
myFormula <- Petal.Length~Species
```

### Step 4: Fit a linear regression model
Call the linreg function to fit a linear regression model.
```{r fit}
linreg_mod <- linreg(myFormula, data=iris)
```

### Step 5: Print the coefficients
Print out the estimated coefficients of the fitted model.
```{r print}
linreg_mod$print()
```

### Step 6: Plot the residuals versus the fitted values
Plot the residuals and the standardized residuals versus the fitted values.
```{r plot}
linreg_mod$plot()
```

### Step 7: Retrieve the residuals
Retrieve the residuals by using *resid()*.
```{r resid}
res <- linreg_mod$resid()
head(res)
```

### Step 8: Retrieve the predicted values
Retrieve the predicted values by using *pred()*.
```{r pred}
pred_val <- linreg_mod$pred()
head(pred_val)
```

### Step 9: Retrieve the coefficient
Retrieve the estimated coefficients by using *coef()*.
```{r coef}
coefs <- linreg_mod$coef()
head(coefs)
```

### Step 10: Print out a summary of the fitted model
Print out a summary of the fitted model.
```{r summary}
linreg_mod$summary()
```
